---
title: "Task1"
author: "Marius Saeltzer"
date: "25 Februar 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a problem set. 


The tasks will always consist of three types of problems

    1. incomplete or errornous code: find the error
    
    2. commenting task: explain what the code is doing by commenting the code
    
    3. recreation task: write code that comes up with the same result as asked for
    

    

Create an element x which is the identity matrix of the form 4x4.
    
```{r}  

x<-


```

Check dimensions,class and whether it is numeric. Store all results in a vector v.

```{r}  

v<-

```


Make sure x is numeric

```{r}

z<-c("3",2,T,"T")


```

Correct the following code! Explain the error message as a comment!

```{r}

c(1,0,0,0)==x[,5]

``` 

Correct the following code! Explain the error message as a comment!

```{r}

is.nA(x)

``` 


Correct the following code! Explain the error message as a comment!

```{r}

paste("There are about 5",*5,"students in this course.")

``` 
