---
title: "Task1"
author: "Marius Saeltzer"
date: "25 Februar 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a problem set. We start in course where your group members and me can help you if necessary.

This is an rmd file. It is important that the package "knitr" is always installed. If you think you are done, click "knit" (the little blue wool ball).

If the task is done correctly and all errors are eliminated, the computer turns it into a nice looking html you can compare in your group. If you come up with different results, you can ask me in the subsequent session.

The tasks will always consist of three types of problems

    1. incomplete or errornous code: find the error
    
    2. commenting task: explain what the code is doing by commenting the code
    
    3. recreation task: write code that comes up with the same result as asked for
    
```{r}
load("solution.Rdata")

```


    

Create an element x which is the identity matrix of the form 4x4.
    
```{r}  

x<-


## Solution Checker
if(identical(solution[[1]],x)==F){stop()}

```

Check dimensions,class and whether it is numeric. Store all results in a vector v.

```{r}  

v<-

## Solution Checker
if(identical(solution[[2]],v)==F){stop()}

```


Make sure x is numeric

```{r}

z<-c("3",2,T,"T")

if(is.numeric(z)==F){stop()}

```

Correct the following code! Explain the error message as a comment!

```{r}

c(1,0,0,0)==x[,5]

``` 

Correct the following code! Explain the error message as a comment!

```{r}

is.nA(x)

``` 


Correct the following code! Explain the error message as a comment!

```{r}

paste("There are about 5",*5,"students in this course.")

``` 
